Beyond conceptual eco-physiological hypotheses: insights from a general size and trait-based model of ectotherm thermal performance

### 2018-03-21 draft_model_only.Rnw
lots of changes in response to Ken's comments

Here, we propose a general quantitative size- and trait-based eco-physiological model to derive general predictions about temperature impacts on ectotherm physiology, performance and ecology. We describe simple size-dependent physiological processes within an ecological context, and, using a simple optimisation argument, show that observed ecological responses of different life-history strategies can be predicted on the basis of optimised bio-energetics under different temperatures.

@@

where $f(w,T)$ is the feeding level ([0,1]) at weight $w$ and temperature $T$, determined by the fraction of time spent foraging (or proportion of maximum attack rate) $\tau$, consumption rate $\gamma w^p \Theta$ (search rate $\gamma w^p$ times prey availability $\Theta$) and maximum consumption $h w^q$. In the following, we will refer to $\tau$ as the activity fraction for sake of generality. Maximum consumption, determined by digestive (enzymatic) processes (\cite{jeschke_predator_2002, sentis_parsing_2013}), is assumed to scale with temperature as $c_T = e^{E_a(T-T_0)/kTT_0}$. Available energy $P(w,T)$ is determined by supply ($S(w,T)$) from prey consumption ($f(w,T) h c_T w^q$), with $\beta$ a loss due to specific dynamic action (SDA, or heat increment; the energy spent absorbing food), and $\phi$ is the fraction of food excreted and egested. Metabolic demands ($D(w,T)$) are those of standard metabolism ($k w^n$), as well as active metabolism, scaled in units of standard metabolism as $\delta k w$, with $\delta$ the factorial aerobic scope (FAS). The activity fraction $\tau$ determines the fraction of time that the active metabolism cost applies. Note that active metabolism scales directly with $w$, owing to muscular demands scaling approximately isometrically with weight (\cite{glazier_activity_2009,brett1965relation}).

@@

Temperature affects growth via its effects on the energy budget and the investment of available (surplus) energy into reproduction and growth. The change in allocation to reproduction with size and age in variable environmental conditions is described by the maturation reaction norm, which is generally defined as the probability of maturing at a certain age under different growth conditions (\cite[e.g., ][]{dieckmann_probabilistic_2007}). We used a slightly different definition here since our model is not probabilistic at the population level: we defined the reaction norm as the mid-point of a logistic allocation function that determines investment in reproduction as a function of age and size. The allocation was parametrised as 
$$\phi(z) = 1/(1+\exp(-(c*z))) \text{, where }$$
$$z(t, w_t) = (w_t-w^*)\times \text{cos}(\text{atan}(b))-t\times \text{sin}(\text{atan}(b))$$ 
rotates the coordinate system about the slope $b$ of the reaction norm, $t$ is the age, $w$ is the mass, $w^*$ is the intercept of the reaction norm, and $c$ determines how rapidly energy allocation shifts from somatic growth to reproduction.

The slope of maturation reaction norms is evolutionarily determined by the strength of the co-variation between growth and mortality for a given population in a given environment. Strongly positive co-variation leads to the relatively flat reaction norms observed for most fish populations (\cite{marty_impact_2011}). This covariation is probably the consequence of good growth conditions (e.g., from increased prey density) altering the baseline mortality $\rho$ and the risk of foraging $\eta$ (e.g., by attracting predators). We did not explicitly model these interactions here (aside from the dependence of mortality on $\tau$), but rather assumed that reaction norms evolved over regimes of relatively stable temperature and growth variation in the past. Consequently, we assumed that the evolved slope of the reaction norm is a fixed trait over the time-scales considered here for a particular species or population, and for simplicity and generality we assume a flat reaction norm \ref{fig:MRN}). The intercept of this reaction norm was found numerically by maximising fitness ($R0$, see below) at the reference temperature for our simulations (15\degree C), and this intercept was asumed fixed as temperatures change.

@@

To ensure a level of generality, we explored ecological impacts of optimised behaviour at different temperatures in a trait-based context. Specifically, we aimed to contrast species along a gradient of life-history that, at the one end, maximises production (energy acquisition; henceforth called the P-strategy) at the cost of increased metabolism and mortality, and at the opposing end minimises mortality and metabolic costs at the expense of production (henceforth M-strategy). This axis leads to an approximately constant ratio of production to mortality, and corresponds to a line of equal size in the life-history space proposed by \citeauthor{charnov_evolutionary_2013} (\citeyear{charnov_evolutionary_2013}). In other words, this axis contrasts species of similar size (here $W_{\infty}\sim 30$cm) with defensive/sluggish versus active life-histories.

To implement this axis, we used the result that species with a more
active, production oriented life-history (e.g., predatory pelagic
fish) have a higher standard metabolism and lower weight scaling of
metabolic costs
(\cite{priede_metabolic_1985,killen_intraspecific_2010}). We assumed
that higher standard metabolism is due to increased digestive capacity
(i.e., is used for gut maintenance), though high muscle mass and a
larger heart will also contribute to higher standard metabolism in
active species (\cite{priede_metabolic_1985}). In practice, we assumed
that approximately 50\% of the standard metabolic cost is due to gut
maintenance, such that a doubling of the maximum ingestion leads to a
50\% increase in standard metabolic cost (i.e., $\Delta_k/\Delta_h=0.5$). We further assumed that such active species have a less effective refuge from predators and therefore have a higher constant mortality, but lower mortality related to activity (i.e., $M^{M-strat}(w) = (0.1+6\tau) w^{q-1}$ and $M^{M-strat}(w) = (1+\tau) w^{q-1}$). Exact parameter values for these trait scenarios are given in Table \ref{tab:parameters}.

We further contrasted species with oxygen limitation at high temperatures (i.e., species with a uni-modal metabolic scope) with species that do not experience oxygen limitation at high temperatures (at least not up to a lethal temeprature, where death may be induced by sudden failure to deliver oxgyen to vital organs). In practice, this was achieved as described above by setting the maximum oxygen deliver close to the lethal temperature (\ref{fig:O2_fig}).

\begin{table}
\caption{Parameters of the constrained activity model for two scenarios: M-strategy and P-strategy species. For parameters with dual values (i.e., x/y), the former reflects species with a domed maximum oxygen supply (MOS) with respect to temperature, whereas the latter corresponds to species with a continuously increasing MOS.}
\label{tab:parameters}
\begin{tabular}{llcc}
 \multirow{2}{*}{Description} &  \multirow{2}{*}{Symbol (unit)} & \multicolumn{2}{r}{Value} \\
 \cmidrule{3-4}
 & & M-strategy  & P-strategy \\

@@

\addlinespace
\multicolumn{4}{c}{\textbf{Feeding ecology}}\\
Coeff. for Encountered food & $\gamma\Theta$ ($y^{-1}$)& \multicolumn{2}{c}{60} \\
Exponent for clearance rate $\gamma$ & $p$ & \multicolumn{2}{c}{0.8}  \\
Coeff. for Maximum consumption rate & $h$ ($y^{-1}$) & 30 & 60  \\
Exponent for max.~consumption $\gamma$ & $q$ & \multicolumn{2}{c}{0.8} \\
Coeff. for constant mortality & $M$ ($g\cdot y^{-1}$) & 0.1  & 1  \\
Coeff. for activity related mortality & $\rho$ ($y^{-1}$) & 6 & 1\\
Exponent for mortality & $\nu$ & \multicolumn{2}{c}{0.2} \\

@@

Slope & $b$ & \multicolumn{2}{c}{0}\\
Reaction & $c$ & \multicolumn{2}{c}{0.5} \\

@@

source('model.R')

@@

int_m <- 500
int_t <- 500
int_temp <- 50

# Sluggish small
ss <- list(c=0.3,
           beta=0.15,
           delta=4,
           Ea=0.52, 
           gamma=60, 
           h=30, 
           k=1,
           lO=0.5, 
           m=10, 
           n=0.88, 
           O2crit=2, 
           omega=0.25, 
           p=0.8,
           P50=4.2, 
           phi=0.25, 
           q=0.8,
           shape=3,
           temp=seq(5,26,l=int_temp),
           Topt=20, 
           temp_ref=15,
           slope=0,
           tmax=20,
           tr=0.5,
           M=0.1,
           v=6,
           lm=int_m,
           nu=0.8-1,
           dt=int_t,
           min_m=-2,
           max_m= 3
)

# fast small0.8-1
fs <- ss
fs$lO <- 1
fs$k <- 1.5
fs$h <- 60
fs$v <- 1
fs$M <- 1
fs$nu<- 0.8-1
fs$n <- 0.75
fs$p <- 0.8
fs$q <- 0.8
fs$tmax <- 10

@@

scs <- list(ss,fs,sl,fl)
scenarios <- bind_cols(data_frame(Scenario = as.factor(rep(c('M strategy','P strategy','M strategy','P strategy'),each=int_temp)),
                                  O2lim = as.factor(rep(c('Domed','Rising'),each=int_temp*2))),
                       bind_rows(lapply(scs, plot_data_temp)))

@@

Increasing metabolic demands with temperature lead species to adjust activity levels in order to optimise energy gains relative to mortality risk (Figure \ref{fig:activity}). This adjustment is especially pronounced in M-strategy species, for which the over-all activity level is markedly lower and which show a consistent increase in activity over-all sizes for the simulated life-history (Figure \ref{fig:size-act}). A similar increase in activity is observed for small P-strategy individuals (e.g., post-larval) for which even initial activity levels are very high (Figure \ref{fig:size-act}). For these individuals, the increasing activity and metabolic demands leads to an active metabolic rate that is close to their MMR. For all other sizes across the two trait scenarios, oxygen is only limiting to activity at the extremes of the simulated temperature range (Figure \ref{fig:activity}), and only for species with a dome-shaped MOS with respect to temperature. However, larger P-strategy individuals are predicted to show a dome-shaped adjustment in activity levels at intermediate sizes, and a decreasing activity level in response to temperature at large sizes, despite available aerobic scope for activity.

\begin{figure}
\centering
<<activity, fig.height=7,fig.width=6>>=

@@

lab1 <- as_labeller(c(`M strategy Rising` ='c',
                   `M strategy Domed`  ='a',
                   `P strategy Rising` ='g',
                   `P strategy Domed` = 'e'))

lab2 <- as_labeller(c(`M strategy Rising` ='d',
                   `M strategy Domed`  ='b',
                   `P strategy Rising` ='h',
                   `P strategy Domed` = 'f'))

ggsc <- ggplot(sc_act %>% filter(Dispatch=='M strategy Domed')) + 
  geom_line(aes(x=Temperature, y=value, linetype=Activity,col=Activity)) +
  theme_cowplot()+
  scale_color_manual(values = cols)+
  scale_linetype()+
  #facet_wrap(~Dispatch,ncol = 1,labeller = lab1)+
  theme(strip.text = element_text(hjust = 0),
        strip.background = element_blank(),
        panel.background = element_blank(),
        plot.background = element_blank()) + 
  labs(x = expression("Mean temperature " ( degree*C)),
       y=expression('Activity ' (tau)))

@@

ggso <- sco  %>% filter(Dispatch=='M strategy Domed') %>%
  ggplot() + 
   geom_line(aes(x=Temperature, y=value, linetype=Metabolism,col=Metabolism)) +
  theme_cowplot()+
  scale_color_manual(values = cols)+
  scale_linetype()+
  #facet_wrap(~Dispatch,ncol = 1,labeller = lab2)+
  theme(strip.text = element_text(hjust = 0),
        strip.background = element_blank(),
        panel.background = element_blank(),
        plot.background = element_blank()) + 
  labs(x = expression("Mean temperature " ( degree*C)),
       y=expression(Oxygen~(mg.yr^-1)))

@@

@
\caption{Optimum (red long-dashed), maximum (green short-dashed) and realised (blue solid lines) activity levels (left column [a,c,e]) at increasing temperatures for a 10g fish, with corresponding oxygen demand (right column [b,d,f]); maximum oxygen supply (MOS; green short-dashed), standard metabolism (red long-dashed) and realised (active; blue solid lines) metabolic demand, as well as metabolic scope (orange long-dashed) at activity level $\tau$, for M-strategy (slow life history; a--d) and P-strategy (fast life-history; e--h). For each strategy, the first row (i.e., a--b; e--f) depicts a life-history with a dome-shaped MOS for increasing temperature, and the second row (i.e., c--d; g--h) corresponds to a species with continuously increasing MOS with temperature.}
\label{fig:activity}
\end{figure}

<<size-act,fig.cap='Optimum (red long-dashed), maximum (green short-dashed) and realised (blue solid lines) activity levels (left column) at increasing temperatures for fish of 5cm (1.25g; a--b); 10cm (10g; c--d) and 20cm (80g; e--f ), for M-strategy (slow life history; left column [a,c,e]) and P-strategy (fast life-history; right column [b,d,f]) species with a dome-shaped maximum metabolic rate with respect to temperature.',fig.align='center'>>=

@@

ggplot(sc_acts %>% filter(Dispatch=='P strategy Large')) + 
  geom_line(aes(x=Temperature, y=value, linetype=Activity,col=Activity)) +
  theme_cowplot()+
  scale_color_manual(values = cols)+
  scale_linetype()+
  #facet_wrap(~Dispatch,ncol = 2,labeller = laba)+
  theme(strip.text = element_text(hjust = 0),
        strip.background = element_blank(),
        panel.background = element_blank(),
        plot.background = element_blank()) + 
  labs(x = expression("Mean temperature " ( degree*C)),
       y=expression('Activity ' (tau)))

@@

Adjustments to the activity level lead to substantial changes in performance related metrics in both trait scenarios (Figure \ref{fig:feeding}). For M-strategy species, the increases in activity levels lead to relatively stable feeding levels, but a substantial increase in mortality coupled with a slow increase in available energy lead to an over-all decline in the ratio of $P$ to $M$. Available energy shows a dome-shaped response to temperature in M-strategists, and is maximised at relatively high temperatures. However, it is limited by oxygen availability only at high temperatures in species with a dome-shaped MOS. Growth efficiency ($P/f$) follows a near identical trend due to the relatively flat response in $f$.

@@

<<feeding,fig.cap='Feeding level (blue solid line), available energy $P$ (green long-dashed), efficiency (orange dotted), mortality (red dotted-dashed) and the ratio of $P$ to $M$ (yellow solid) are impacted by changing activity and the metabolic response to temperature for a growing fish at 10cm length (10g). Responses are shown for M- and P-strategy (top and bottom rows, respectively), for species with and without oxygen limitation (left and right columns, respectively). Energy and fitness are plotted relative to maximum over all temperatures.',fig.align='center'>>=

lab <- as_labeller(c(`M strategy Rising` ='b',
                   `M strategy Domed`  ='a',
                   `P strategy Rising` ='d',
                   `P strategy Domed` = 'c'))

sc_feed <- scenarios %>%
  #filter(!grepl('large',Scenario)) %>% 
  group_by(Scenario,O2lim) %>%
  mutate(Consumption=sc(Consumption),
         Efficiency=sc(Efficiency),
         Feeding=`Feeding level`,
         `Available Energy` = sc(`C for growth`),
         `Energy` = ifelse(`Available Energy`<0,0,`Available Energy`),
         Fitness = sc(Energy/M)) %>% 
  tidyr::gather(Rate,value,Feeding,Energy,Efficiency,M,Fitness) %>%
  mutate(Rate = factor(Rate,levels = c('Feeding','Energy','Efficiency','M','Fitness')),
         Dispatch = paste(Scenario,O2lim))

ggplot(sc_feed%>% filter(Dispatch=='M strategy Domed')) + 
  geom_line(aes(x=Temperature, y=value, linetype=Rate,col=Rate,group=Rate)) +
  theme_cowplot()+
  #facet_wrap(~Dispatch,labeller = lab,ncol=2)+
  theme(strip.text = element_text(hjust = 0),
        strip.background = element_blank(),
        panel.background = element_blank(),
        plot.background = element_blank())+
  scale_linetype_manual(values=c(1:4,1))+
  scale_color_manual(values = cols)+
  labs(x = expression("Mean temperature " ( degree*C)),
       y='Relative rate')

@@

Simulated growth curves illustrate the ontogenetic consequences of increasing temperatures (Figure \ref{fig:norms}). For all trait scenarios, fastest growth occurred at temperatures close to the lethal temperature, with declining growth for oxygen limited species at high temperatures. This can be explained by ontogenetic shifts in temperature optima for growth (Figure \ref{fig:size-rates}): for small individuals, available energy and growth efficiency consistently peak at high temperatures, but this peak rapidly moves to lower temperatures as individuals for either trait-scenario grow. For large individuals, growth is optimised at relatively low temperatures, leading to larger asymptotic size at lower temperatures.

>>>>

Here, we propose a mechanistic, quantitative size- and trait-based eco-physiological model to derive general predictions about temperature impacts on ectotherm physiology, performance and ecology. Our aim was to produce a model that is general --- a model that incorporates simple descriptions of general processes, but ignoring particularities that may apply to specific species and contexts. In other words, it is the model framework that is general, rather than the particular parameters and the responses they induce. We show that this simple framework can lead to rather different temperature responses that depend on traits, organism size and ecological factors such as food availibility and mortality.

@@

where $f(w,T)$ is the feeding level ([0,1]) at weight $w$ and temperature $T$, determined by consumption rate $\gamma w^p \Theta$ (search rate $\gamma w^p$ times prey availability $\Theta$), maximum consumption $h w^q$, and $\tau$, which can be interpreted as the proportion of time spent foraging (or proportion of maximum attack rate) or a multiplyer on a basal foraging speed, such that at $\tau=1$, foraging occurs at maximum speed. In the following, we will refer to $\tau$ as the activity fraction for sake of generality. Maximum consumption, determined by digestive (enzymatic) processes (\cite{jeschke_predator_2002, sentis_parsing_2013}), is assumed to scale with temperature according to the Arrhenius scaling $c_T = e^{E_a(T-T_0)/k \times T \times T_0}$. Available energy $P(w,T)$ is determined by supply ($S(w,T)$) from prey consumption ($f(w,T) h c_T w^q$), with $\beta$ a loss due to specific dynamic action (SDA, or heat increment; the energy spent absorbing food), and $\phi$ is the fraction of food excreted and egested. Metabolic demands ($D(w,T)$) are those of standard metabolism ($k w^n$), as well as active metabolism, scaled in units of standard metabolism as $\delta k w$, with $\delta$ the factorial aerobic scope (FAS). The activity fraction $\tau$ determines the proportion of total metabolic costs that are due to active metabolism. Note that we assume that active metabolism scales directly with $w$, owing to muscular demands scaling approximately isometrically with weight (\cite{glazier_activity_2009,brett1965relation}). In practice, the scaling may not be strictly isometric owing to reduced drag at larger size (\cite[e.g., ][]{ware_bioenergetics_1978}), as well as non-strictly isometric growth of muscle mass with weight. However, the precise exponents of standard and foraging metabolism only alter predictions quantitatively, but do not change the over-all results.

@@

Temperature affects growth via its effects on the energy budget and the investment of available (surplus) energy into reproduction and growth. The change in allocation to reproduction with size and age in variable environmental conditions is described by the maturation reaction norm, which is generally defined as the probability of maturing at a certain age under different growth conditions (\cite[e.g., ][]{dieckmann_probabilistic_2007}). We defined the reaction norm as the mid-point of a logistic allocation function that determines investment in reproduction as a function of age and size. The allocation was parametrised as 
$$\phi(z) = 1/(1+\exp(-(c*(w_t-w^*)))),$$$$\phi(z) = 1/(1+\exp(-(c*z))) \text{, where }$$
$$z(t, w_t) = (w_t-w^*)\times \text{cos}(\text{atan}(b))-t\times \text{sin}(\text{atan}(b))$$  
rotates the coordinate system about the slope $b$ of the reaction norm, $t$ is the age, $w$ is the mass, $w^*$ is the intercept of the reaction norm, and $c$ determines how rapidly energy allocation shifts from somatic growth to reproduction.

The slope of maturation reaction norms is evolutionarily determined by the strength of the co-variation between growth and mortality for a given population in a given environment. Strongly positive co-variation leads to the relatively flat reaction norms observed for most fish populations (\cite{marty_impact_2011}). This covariation is probably the consequence of good growth conditions (e.g., from increased prey density) altering the baseline mortality $\rho$ and the risk of foraging $\eta$ (e.g., by attracting predators). We did not explicitly model these interactions here (aside from the dependence of mortality on $\tau$), but rather assumed that reaction norms evolved over regimes of relatively stable temperature and growth variation in the past. Consequently, we assumed that the evolved slope of the reaction norm is a fixed trait over the time-scales considered here for a particular species or population, and for simplicity and generality we assume a slightly negative slope in the reaction norm \ref{fig:MRN}). The intercept of this reaction norm was found numerically by maximising fitness ($R0$, see below) at the reference temperature for our simulations (15\degree C), and this intercept was asumed fixed as temperatures change.

As growth is also fundamentally driven by resource availibility, we contrast the growth reponse to temperature at the baseline level (Table \ref{tab:parameters}), with a 30\% reduction and increase in available resources.

@@

To ensure a level of generality, we explored ecological impacts of optimised behaviour at different temperatures in a trait-based context. Specifically, we aimed to contrast species along a gradient of life-history that, at the one end, maximises production (energy acquisition; henceforth called the P-strategy) at the cost of increased metabolism and mortality, and at the opposing end minimises mortality and metabolic costs at the expense of production (henceforth M-strategy). This axis leads to an approximately constant ratio of production to mortality, and corresponds to a line of equal size in the life-history space proposed by \citeauthor{charnov_evolutionary_2013} (\citeyear{charnov_evolutionary_2013}). In other words, this axis contrasts species of similar size (here $L_{\infty}\sim 30$cm) with defensive/sluggish versus active life-histories.

To implement this axis, we used the result that species with a more
active, production oriented life-history (e.g., predatory pelagic
fish) have a higher standard metabolism and lower weight scaling of
metabolic costs
(\cite{priede_metabolic_1985,killen_intraspecific_2010}). We assumed
that higher standard metabolism is due to increased digestive capacity
(i.e., is used for gut maintenance), though high muscle mass and a
larger heart will also contribute to higher standard metabolism in
active species (\cite{priede_metabolic_1985}). In practice, we assumed
that approximately 50\% of the standard metabolic cost is due to gut
maintenance, such that a doubling of the maximum ingestion leads to a
50\% increase in standard metabolic cost. We further assumed that such active species have a less effective refuge from predators and therefore have a higher constant mortality, but lower mortality related to activity (i.e., $M^{M-strat}(w) = (0.1+6\tau) w^{q-1}$ and $M^{M-strat}(w) = (1+\tau) w^{q-1}$). Exact parameter values for these trait scenarios are given in Table \ref{tab:parameters}. Together, these assumed trait differences lead to very different recological and bio-energetic responses of M- and P-strategists (Figure \ref{fig:scenarios}), with $\tau_{\text{opt}}$ found at lower activity levels for M-strategists, as high activity induces exceedingly high mortality and decreasing energy efficiency (i.e., available energy relative to feeding level) at high activity levels. A slower increase in available energy and M with $tau$ for P-strategists leads to a higher $\tau_{\text{opt}}$.

We further contrasted species with oxygen limitation at high temperatures (i.e., species with a uni-modal metabolic scope) with species that do not experience oxygen limitation at high temperatures (at least not up to a lethal temeprature, where death may be induced by sudden failure to deliver oxgyen to vital organs). In practice, this was achieved as described above by setting the maximum oxygen deliver close to the lethal temperature (Figure \ref{fig:O2_fig}).

\begin{table}
\caption{Parameters of the constrained activity model for two scenarios: M-strategy and P-strategy species. For parameters with dual values (i.e., x/y), the former reflects species with a domed maximum oxygen supply (MOS) with respect to temperature, whereas the latter corresponds to species with a continuously increasing MOS. Values in brackets for $\\gamma$ are alternative resource availibility scenarios.}
\label{tab:parameters}
\begin{tabular}{llcc}
 \multirow{2}{*}{Description} &  \multirow{2}{*}{Symbol (unit)} & \multicolumn{2}{r}{Value} \\
 \cmidrule{3-4}
 & & M-strategy  & P-strategy \\

@@

\addlinespace
\multicolumn{4}{c}{\textbf{Feeding ecology}}\\
Coeff. for Encountered food & $\gamma\Theta$ ($y^{-1}$)& \multicolumn{2}{c}{60 (40/80)} \\
Exponent for clearance rate $\gamma$ & $p$ & \multicolumn{2}{c}{0.8}  \\
Coeff. for Maximum consumption rate & $h$ ($y^{-1}$) & 30 & 60  \\
Exponent for max.~consumption $\gamma$ & $q$ & \multicolumn{2}{c}{0.8} \\
Coeff. for constant mortality & $M$ ($g\cdot y^{-1}$) & 0.1  & 1  \\
Coeff. for activity related mortality & $\rho$ ($y^{-1}$) & 6 & 1\\
Exponent for mortality & $\nu$ & \multicolumn{2}{c}{0.2} \\

@@

Slope &  & \multicolumn{2}{c}{0}\\
Reaction & $c$ & \multicolumn{2}{c}{0.5} \\

@@

source('model.R')
cols <- c('#4477AA','#44AA77','#AA7744','#AA4488','#DDDD77')

@@

int_m <- 100
int_t <- 100
int_temp <- 30

# Sluggish small
ss <- list(c=0.3,
           beta=0.15,
           delta=4,
           Ea=0.52, 
           gamma=60, 
           h=30, 
           k=1,
           lO=0.5, 
           m=10, 
           n=0.88, 
           o=1,
           O2crit=2, 
           omega=0.25, 
           p=0.8,
           P50=4.2, 
           phi=0.25, 
           q=0.8,
           shape=3,
           temp=seq(5,26,l=int_temp),
           Topt=20, 
           temp_ref=15,
           slope=-0.5,
           tmax=20,
           tr=0.5,
           M=0.1,
           v=6,
           lm=int_m,
           nu=0.8-1,
           dt=int_t,
           min_m=-2,
           max_m= 3
)

# fast small0.8-1
fs <- ss
fs$lO <- 1
fs$k <- 1.5
fs$h <- 60
fs$v <- 1
fs$M <- 1
fs$nu<- 0.8-1
fs$n <- 0.75
fs$o <- 1
fs$p <- 0.8
fs$q <- 0.8
fs$tmax <- 10

@@

scs <- list(ss,fs,sl,fl)

baseline <- bind_cols(data_frame(Scenario = as.factor(rep(c('M strategy','P strategy','M strategy','P strategy'),each=int_temp)),
                                  O2lim = as.factor(rep(c('Domed','Rising'),each=int_temp*2))),
                       bind_rows(lapply(scs, get_model_out,seq(0.05,1,l=int_temp))))

scenarios <- bind_cols(data_frame(Scenario = as.factor(rep(c('M strategy','P strategy','M strategy','P strategy'),each=int_temp)),
                                  O2lim = as.factor(rep(c('Domed','Rising'),each=int_temp*2))),
                       bind_rows(lapply(scs, plot_data_temp)))

@@

Increasing metabolic demands with temperature lead species to adjust activity levels in order to optimise energy gains relative to mortality risk (Figure \ref{fig:activity}). This adjustment is especially pronounced in M-strategy species, for which the over-all activity level is markedly lower and which show a consistent increase in activity over-all sizes for the simulated life-history (Figure \ref{fig:size-act}). A similar increase in activity is observed for small P-strategy individuals (e.g., post-larval) for which even initial activity levels are very high (Figure \ref{fig:size-act}). For these individuals, the increasing activity and metabolic demands lead to an active metabolic rate that is close to their MMR. For all other sizes across the two trait scenarios, oxygen is only limiting to activity at the extremes of the simulated temperature range (Figure \ref{fig:activity}), and only for species with a dome-shaped MOS with respect to temperature. For species with a rising MOS with temperature, oxygen is not limiting (Figure \ref{fig:activitys}). However, larger P-strategy individuals are predicted to show a dome-shaped adjustment in activity levels at intermediate sizes, and a decreasing activity level in response to temperature at large sizes, despite available aerobic scope for activity.

<<scenarios,fig.cap='Feeding level (blue solid line), available energy $P$ (green long-dashed), efficiency (orange dotted), mortality (red dotted-dashed) and the ratio of $P$ to $M$ (yellow solid) are impacted as a function of changing activity at constant temperature ($T=15$\\degree C) for a growing fish ($L_{\\infty}\\sim 30$cm) at 10cm length (10g). Responses are shown for M- and P-strategy (a and b, respectively), for species with and without oxygen limitation (left and right columns, respectively). Energy and fitness are plotted relative to maximum over all temperatures.',fig.align='center',fig.height=3,fig.width=6>>=

lab <- as_labeller(c(`M strategy Domed`  ='a',
                   `P strategy Domed` = 'b'))

sc_bl <- baseline %>% filter(O2lim=='Domed') %>%
  #filter(!grepl('large',Scenario)) %>% 
  group_by(Scenario,O2lim) %>%
  mutate(Consumption=sc(Consumption),
         Efficiency=sc(Efficiency),
         Feeding=`Feeding level`,
         `Available Energy` = sc(`C for growth`),
         `Energy` = ifelse(`Available Energy`<0,0,`Available Energy`),
         Fitness = sc(Energy/M)) %>% 
  tidyr::gather(Rate,value,Feeding,Energy,Efficiency,M,Fitness) %>%
  mutate(Rate = factor(Rate,levels = c('Feeding','Energy','Efficiency','M','Fitness')),
         Dispatch = paste(Scenario,O2lim))

ggplot(sc_bl) + 
  geom_line(aes(x=tau, y=value, linetype=Rate,col=Rate,group=Rate)) +
  theme_cowplot()+
  facet_wrap(~Dispatch,labeller = lab,ncol=2)+
  theme(strip.text = element_text(hjust = 0),
        strip.background = element_blank(),
        panel.background = element_blank(),
        plot.background = element_blank())+
  scale_linetype_manual(values=c(1:4,1),guide='none')+
  scale_color_manual(values = cols,guide='none')+
   labs(x = expression('Activity ' (tau)),
       y='Relative rate')

\begin{figure}
\centering
<<activity, fig.height=4,fig.width=6>>=

@@

lab1 <- as_labeller(c(`M strategy Domed`  ='a',
                   `P strategy Domed` = 'c'))

lab2 <- as_labeller(c(`M strategy Domed`  ='b',
                   `P strategy Domed` = 'd'))

ggsc <- ggplot(sc_act %>% filter(O2lim=='Domed')) + 
  geom_line(aes(x=Temperature, y=value, linetype=Activity,col=Activity)) +
  theme_cowplot()+
  scale_color_manual(values = cols,guide='none')+
  scale_linetype(guide='none')+
  facet_wrap(~Dispatch,ncol = 1,labeller = lab1)+
  theme(strip.text = element_text(hjust = 0),
        strip.background = element_blank(),
        panel.background = element_blank(),
        plot.background = element_blank()) + 
  labs(x = expression("Mean temperature " ( degree*C)),
       y=expression('Activity ' (tau)))

@@

ggso <- sco %>% filter(O2lim=='Domed') %>%
  ggplot() + 
   geom_line(aes(x=Temperature, y=value, linetype=Metabolism,col=Metabolism)) +
  theme_cowplot() +
  scale_color_manual(values = cols,guide='none')+
  scale_linetype(guide='none')+
  facet_wrap(~Dispatch,ncol = 1,labeller = lab2)+
  theme(strip.text = element_text(hjust = 0),
        strip.background = element_blank(),
        panel.background = element_blank(),
        plot.background = element_blank()) + 
  labs(x = expression("Mean temperature " ( degree*C)),
       y=expression(Oxygen~(mg.yr^-1)))

@@

@
\caption{Optimum (red long-dashed), maximum (green short-dashed) and realised (blue solid lines) activity levels (left column [a,c]) at increasing temperatures for a 10g fish with a dome-shaped MOS with increasing temperature (maximum at 20\degree C), with corresponding oxygen demand (right column [b,d]); maximum oxygen supply (MOS; green short-dashed), standard metabolism (red long-dashed) and realised (active; blue solid lines) metabolic demand, as well as metabolic scope (orange long-dashed) at activity level $\tau$, for M-strategy (slow life history; aandb) and P-strategy (fast life-history; candd).}
\label{fig:activity}
\end{figure}

<<size-act,fig.cap='Optimum (red long-dashed), maximum (green short-dashed) and realised (blue solid lines) activity levels (left column) at increasing temperatures through ontogeny for a for fish with a $L_{\\infty}\\sim 30$cm at 1.25g (5cm; a--b); 10g (10cm; c--d) and 80g (20cm; e--f), for M-strategy (slow life history; left column [a,c,e]) and P-strategy (fast life-history; right column [b,d,f]) species with a dome-shaped maximum metabolic rate with respect to temperature.',fig.align='center'>>=

@@

ggplot(sc_acts) + 
  geom_line(aes(x=Temperature, y=value, linetype=Activity,col=Activity)) +
  theme_cowplot()+
  scale_color_manual(values = cols,guide='none')+
  scale_linetype(guide='none')+
  facet_wrap(~Dispatch,ncol = 2,labeller = laba)+
  theme(strip.text = element_text(hjust = 0),
        strip.background = element_blank(),
        panel.background = element_blank(),
        plot.background = element_blank()) + 
  labs(x = expression("Mean temperature " ( degree*C)),
       y=expression('Activity ' (tau)))

@@

Temeprature and metabolic demand-driven adjustments to the activity level lead to substantial changes in performance related metrics in both trait scenarios (Figure \ref{fig:feeding}). For M-strategy species, the increases in activity levels lead to relatively stable feeding levels, but a substantial increase in mortality coupled with a slow increase in available energy leads to an over-all decline in the ratio of $P$ to $M$. Available energy shows a dome-shaped response to temperature in M-strategists, and is maximised at relatively high temperatures. However, it is limited by oxygen availability only at high temperatures in species with a dome-shaped MOS. Growth efficiency ($P/f$) follows a near identical trend due to the relatively flat response in $f$.

@@

<<feeding,fig.cap='Feeding level (blue solid line), available energy $P$ (green long-dashed), efficiency (orange dotted), mortality (red dotted-dashed) and the ratio of $P$ to $M$ (yellow solid) are impacted by changing activity and the metabolic response to temperature for a growing fish ($L_{\\infty}\\sim 30$cm) at 10cm length (10g). Responses are shown for M- and P-strategy (a and b, respectively), for species with and without oxygen limitation (left and right columns, respectively). Energy and fitness are plotted relative to maximum over all temperatures.',fig.align='center',fig.height=3,fig.width=6>>=

lab <- as_labeller(c(`M strategy Domed`  ='a',
                   `P strategy Domed` = 'b'))

sc_feed <- scenarios %>% filter(O2lim=='Domed') %>%
  #filter(!grepl('large',Scenario)) %>% 
  group_by(Scenario,O2lim) %>%
  mutate(Consumption=sc(Consumption),
         Efficiency=sc(Efficiency),
         Feeding=`Feeding level`,
         `Available Energy` = sc(`C for growth`),
         `Energy` = ifelse(`Available Energy`<0,0,`Available Energy`),
         Fitness = sc(Energy/M)) %>% 
  tidyr::gather(Rate,value,Feeding,Energy,Efficiency,M,Fitness) %>%
  mutate(Rate = factor(Rate,levels = c('Feeding','Energy','Efficiency','M','Fitness')),
         Dispatch = paste(Scenario,O2lim))

ggplot(sc_feed) + 
  geom_line(aes(x=Temperature, y=value, linetype=Rate,col=Rate,group=Rate)) +
  theme_cowplot()+
  facet_wrap(~Dispatch,labeller = lab,ncol=2)+
  theme(strip.text = element_text(hjust = 0),
        strip.background = element_blank(),
        panel.background = element_blank(),
        plot.background = element_blank())+
  scale_linetype_manual(values=c(1:4,1),guide='none')+
  scale_color_manual(values = cols,guide='none')+
  labs(x = expression("Mean temperature " ( degree*C)),
       y='Relative rate')

@@

Simulated growth curves illustrate the ontogenetic consequences of increasing temperatures (Figure \ref{fig:norms}). For both trait scenarios, fastest growth occurred at temperatures close to the lethal temperature, with declining growth for oxygen limited species at high temperatures. This can be explained by ontogenetic shifts in temperature optima for growth (Figure \ref{fig:size-rates}): for small individuals, available energy and growth efficiency consistently peak at high temperatures, but this peak rapidly moves to lower temperatures as individuals for either trait-scenario grow. For large individuals, growth is optimised at relatively low temperatures, leading to larger asymptotic size at lower temperatures. Resource availibility storngly modulates this growth response to temeprature: low resource availibility leads to strong differences in asymptotic size, whereas high food availibility leads to an almost inverted response
